---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r}
tuition = read.csv('tuition-final.csv')
summary(tuition)
head(tuition)
```




2)
```{r}
tuition$Acc.rate = tuition$Accepted/tuition$Applied
head(tuition)
```

3)
```{r}
unc = tuition[tuition$Name == 'University of North Carolina at Chapel Hill',]
unc
```
At line 682



4)
```{r}
add1 <- function(a, b){
  
  # add the variables
  c = a + b
  return(c)
  
}

add2 <- function(a, b = 3){
  
  # add the variables
  c = a + b
  return(c)
  
}

# Try adding 5 and 7
add1(5, 7)
add2(5, 7)

# Try adding one variable
#add1(5)
add2(5)
```
b is assigned 3 unless specified in the function call. Allows the function to be called with just a. Because of this, add2 works with just a as a parameter, while add1 causes an error.





5)
```{r}
beta1 <- function(r, Sx, Sy){
  return(r*(Sy/Sx))
}

beta0 <- function(r, Sx, Sy, ybar, xbar){
  b1 = beta1(r, Sx, Sy)
  b0 = ybar-b1*xbar
  return(b0)
}

beta1(1,2,0)
beta0(1,2,0,3,4)
```
Functions above^



6) Sy = 0 would always result in beta1 being 0. That would mean there is no linear relationship between the two variables.




```{r}
plot(tuition$Avg.SAT, tuition$Out.Tuition, main = "title", xlab = "label", ylab = "label", pch = 7, cex = 2, col = "blue")
```
7)
```{r}
plot(tuition$Avg.SAT, tuition$Out.Tuition, main = "Sat vs Tuition", xlab = "Average Sat Score", ylab = "Tuiton ($)", pch = 1, cex = 0.5, col = "blue")
```
pch changes the shape of the object that is plotted. cex changes the size of the plotted object.




8)
```{r}
plot(tuition$Avg.SAT, tuition$Out.Tuition, main = "Sat vs Tuition", xlab = "Average Sat Score", ylab = "Tuiton ($)", pch = 1, cex = 0.5, col = tuition$Public)
```
If the school is public it is drawn in one color, looks to be black in this case. If a school is not public, it is drawn in another color, which is red here.





9)
```{r}
plot(tuition$Avg.SAT, tuition$Out.Tuition, main = "Sat vs Tuition", xlab = "Average Sat Score", ylab = "Tuiton ($)", pch = 1, cex = 0.5, col = tuition$Public)
abline(0,10)
```





10)
```{r}
tuition_clean = tuition[!is.na(tuition$Avg.SAT ) & !is.na(tuition$Out.Tuition),]

sat = tuition_clean$Avg.SAT
outtuition = tuition_clean$Out.Tuition
SSX = sum((sat-mean(sat))^2)
SSY = sum((outtuition-mean(outtuition))^2)
SXY = sum((sat-mean(sat))*(outtuition-mean(outtuition)))
SX = sqrt(SSX/(length(sat)-1))
SY = sqrt(SSY/(length(sat)-1))
r = SXY/sqrt(SSX*SSY)
b1 = beta1(r, SX, SY)
b0 = beta0(r, SX, SY, mean(outtuition), mean(sat))

b0
b1
beta1(r, sd(tuition_clean$Avg.SAT), sd(tuition_clean$Out.Tuition))

plot(tuition$Avg.SAT, tuition$Out.Tuition, main = "Sat vs Tuition", xlab = "Average Sat Score", ylab = "Tuiton ($)", pch = 1, cex = 0.5, col = tuition$Public)
abline(b0,b1)
```

There appears to be a positive relationship between average SAT score and tuition cost. As the average Sat score increases, the tuition cost of the school increases.






11)
```{r}
predict_yval <- function(X, Y, x_new){
  SSX = sum((X-mean(X, na.rm = T))^2, na.rm = T)
  SSY = sum((Y-mean(Y, na.rm = T))^2, na.rm = T)
  SXY = sum((X-mean(X, na.rm = T))*(Y-mean(Y, na.rm = T)), na.rm = T)
  SX = sqrt(SSX/(length(X)-1))
  SY = sqrt(SSY/(length(X)-1))
  r = SXY/sqrt(SSX*SSY)
  
  b1 = beta1(r, SX, SY)
  b0 = beta0(r, SX, SY, mean(Y, na.rm = T), mean(X, na.rm = T))
  return(b0+x_new*b1)
  
}
```


```{r}
# Find UNC values
unc = tuition[tuition$Name == 'University of North Carolina at Chapel Hill',]
x_unc = unc$Avg.SAT
y_unc = unc$Out.Tuition

# Find Duke values
duke = tuition[tuition$Name == 'Duke University',]
x_duke = duke$Avg.SAT
y_duke = duke$Out.Tuition

y_unc
y_duke
# Predict tuitions
predict_yval(tuition_clean$Avg.SAT, tuition_clean$Out.Tuition, x_unc)
predict_yval(tuition_clean$Avg.SAT, tuition_clean$Out.Tuition, x_duke)
```
The regression overestimates the tuition for UNC by about 5000 dollars and underestimates the tuition for Duke by about 3600 dollars. According to this model, you would be getting a good deal at UNC, but not Duke.




```{r}
pred = lm(outtuition~sat)
-9173.7964 + x_duke*19.7979
```

