---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

1) a)

```{r}
table = read.table('TABLE346.txt', header = T)
x = table$Time
y = table$Weight
n = length(y)
SSX = sum((x-mean(x))^2)
SSY = sum((y-mean(y))^2)
SXY = sum((x-mean(x))*(y-mean(y)))
SSE = SSY - ((SXY)^2)/SSX
beta1 = SXY/SSX
beta0 = mean(y) - beta1*mean(x)
sdhat = sqrt(SSE/(n-2))
beta1
beta0
sdhat
table
```
beta1 hat= 0.5302
beta0 hat= -0.2273
sdhat = 0.9166


b)
  1) The crystals grow on average 0.5302 grams per hour.
```{r}
beta0 + 15*(beta1)
```
  2)After 15 hours, the crystal is predicted to weigh 7.7264 grams.
```{r}
h12 = 10 * (beta0 + 12*beta1)
h24 = 16 * (beta0+24*beta1)
h24-h12
```
  
  3)A 24 hours crystal will make $138.62 more than a 12 hours crystal
  
c)
```{r}
beta0 + beta1*19
beta0 + beta1*25
beta0 + beta1*40
```
Average weight for 19 hours: 9.8474 grams
25 hours: 13.0289
40 hours: can not be estimated using this regression since 40 hours is outside of our data


d)
```{r}
beta1 = 7391396/224617600
beta0 = 532.44 - 11364*beta1
beta1
beta0
```
  a)beta0 hat= 158.4897
    beta1 hat = 0.0329
    muy(x) = 158.4897 + 0.0329x
    
  b) They would be interested in Y(13,000). They want to compare to a car similar to the ones in the table that are driven 13,000 miles rather than the average of all vehicles driven 13,000 miles. Yhat(13,000) is estimated the same as muyhat(13,000)
```{r}
158.4897 + 0.0329*13000
```
  Yhat(13,000) = 586.1897
  
  c) We want to use muy(16000). muy(16000) is the average first year maintenence cost for all vehicles driven 16000 miles.



2) a)
```{r}
exam = read.table("EXAM.txt", header = T)
exam
fin = exam$FIN
mid = exam$MID
fit = lm(fin~mid)
summary(fit)
```

beta0 hat= -6.132
beta1 hat= 1.082
sd hat= 9.093

b)
```{r}
res = resid(fit)
plot(mid, res, xlab = 'Midterm Score', ylab = 'Residuals')
```

```{r}
stdres = rstandard(fit)
plot(mid, stdres, xlab = 'Midterm Score', ylab = 'Standardized Residuals')
```

It seems to be a good fit. The standardized residuals are distributed around -1 to 1 fairly randomly. There is likely an outlier with a standardized residual of above 3 around a midterm of 60.


c)
```{r}
beta0 = -6.132
beta1 = 1.082
sdres = 9.093
predicted = beta0 + beta1*70
predicted 
pnorm(75, predicted, sdres, lower.tail = F)
```
Predicted final score: 69.608
There is a probability of 0.2766 that the student scores higher than a 75 on the final