---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

1)

a) Target population could be all 1/4 liter bottles of that specific brand of cough syrup stored within a reasonable temperature range compared to the study population.

b)Study population could be 1/4 liter cough syrup bottles of that brand randomly selected from each temperature within a reasonable storage range.

c)Seems to be preselected since there is only one bottle for each temperature from 14 to 30 degrees C. 

d)
```{r}
shelf_life = read.table('shelflif.txt', header = T)
life = shelf_life$ShelfLife
temp = shelf_life$Temp
plot(temp, life)
```
It seems reasonable for linear regression. The plot looks linear. There might be traces of a non-linear shape but I wouldn't conclude anything.

e)
NH: beta1 = 0       AH: beta1 =/= 0
```{r}
SSX = 484.5
SSY = 96294.9
SXY = -6663.49
SSE = SSY - (SXY)^2/SSX
sd = sqrt(SSE/16)
beta1 = SXY/SSX
se = sd/sqrt(SSX)
tc = beta1/se
pt(tc, 16)*2
```
P-value is 5.932e-12. Since the p-value < 0.05, we reject the NH at the 95% confidence level and conclude the shelf life is related to temperature.

f) 
```{r}
beta0 = 11341/18 - beta1*387/18
beta0 
beta0 + 0*beta1
```
The estimate for average shelf life at 0 degrees is 925.7522 days though it is an extrapolation. It would not be reasonable to use this estimate.

g)
```{r}
fifteen = beta0 + 15*beta1
t = qt(.975, 16)
xbar = 387/18
yse = sd*sqrt(1/18+(15-xbar)^2/SSX)
lower = fifteen - t*yse
upper = fifteen + t*yse
fifteen
lower
upper
```
Estimate for average self life when stored at 15 degrees is 719.4522 days
C[705.7979 <= muy(15) <= 733.1066] = 0.95


h)
```{r}
lower = beta1 - t*se
upper = beta1 + t*se
lower 
upper
```
C[-15.39514 <= beta1 <= -12.1115] = 0.95

Since 0 is not in the confidence interval and agrees with our hypothesis test in e), we can conclude that there is a relationship between shelf life and storage temperature

i)
NH: muy(13) <= 650        AH: muy(13) > 650
```{r}
thirteen = beta0 + beta1*13
muyse = sd*sqrt(1/18+(13-xbar)^2/SSX)
tc = (thirteen - 650)/muyse
pt(tc, 16, lower.tail = F)
```
Since p-value < 0.05, we reject NH and conclude at a 95% confidence level that the average shelf life for bottles stored at 13 degrees C is at least 650 days


j)
```{r}
lower = thirteen - t*muyse
upper = thirteen + t*muyse
upper 
lower
```
C[730.6094 <= muy(13) <= 763.3084] = 0.95
Since values 650 and less are not in the 95% confidence interval, we can reject NH and conclude that the average shelf life of a bottle stored at 13 deg C is at least 650 days


2)

a)
```{r}
SSR = SSY-SSE
MSR = SSR
MSE = SSE/16
MSY = SSY/17
Fc = MSR/MSE
aov_table = matrix(c(1, SSR, MSR, Fc, 16, SSE, MSE, NA, 17, SSY, MSY, NA), ncol = 4, byrow = T)
colnames(aov_table) = c('DF', 'Sum Sq', 'Mean Sq', 'F value')
rownames(aov_table) = c('Regression', 'Error', 'Total')
aov_table
```

b)
```{r}
pf(Fc, 1, 16, lower.tail = F)
```
P-value is 5.932e-12. The p-value is very close to zero so we likely reject the NH. 

c)
```{r}
tc = beta1*sqrt(SSX)/sd
tc
2*pt(tc, 16)
```
P-value is 5.932e-12. The p-value is very close to zero so we likely reject the NH. 

d)
```{r}
Fc == tc^2
Fc
tc^2
```
F equals tc squared. P-value from both is 5.932e-12.

e)We conclude that beta1 =/= 0  with 99% confidence (could be more or less, p-value very small.

f)
```{r}
se = sd/sqrt(SSX)
t = qt(.995, 16)
lower = beta1 - t*se
upper = beta1 + t*se
lower
upper
```
C[-16.0154 <= beta1 <= -11.4912] = 0.99
Since 0 is not in the interval or even close, we can conclude beta1 is not 0 at a 99% confidence level.

g) In b) and c), we used an f-test and a t-test on the null hypothesis that beta1 equals 0. Both tests provided a p-value of 5.932e-12, which suggests to reject the NH. We showed that Fc equals tc^2 and confirmed both provided the same p-value. In f), we found a 99% confidence interval for beta1 that showed 0 was not in the interval. This further confirmed that beta1 is not 0.


3)
a)
```{r}
exam = read.table('Exam2004.txt', header = T)
fin = exam$FIN
mid = exam$MT
n = nrow(exam)
lm1 = lm(FIN ~ MT, data=exam)
summary(lm(fin~mid))
```
muy(x) hat = 46.6088 + 0.4253x

b)

```{r}
tstar = sqrt(2*qf(0.95,2,n-2))
beta1 = summary(lm1)$coefficients[2]
beta0 = summary(lm1)$coefficients[1]
SSX = sum((mid-mean(mid))^2)
sigma_hat = summary(lm1)$sigma
se = sigma_hat*sqrt(1/n+(mid-mean(mid))^2/SSX)
y1 = beta0 + beta1*mid
y1 = lm1$fitted
upper = y1 + tstar*se
lower = y1 - tstar*se

plot(mid, fin, xlab = 'Midterm', ylab = 'Final')
abline(lm1)
lines(sort(mid), sort(upper), lty=2)
lines(sort(mid), sort(lower), lty=2)
```



c)
```{r}
fit40 = beta0 + beta1*40
fit60 = beta0 + beta1*60
fit80 = beta0 + beta1*80
tstar = qt(0.975, n-2)

interval40 = tstar*sigma_hat*sqrt(1/n + 1 + (40-mean(mid))^2/SSX)

interval60 = tstar*sigma_hat*sqrt(1/n + 1 + (60-mean(mid))^2/SSX)

interval80 = tstar*sigma_hat*sqrt(1/n + 1 + (80-mean(mid))^2/SSX)

c(fit40+interval40, fit40-interval40)
c(fit60+interval60, fit60-interval60)
c(fit80+interval80, fit80-interval80)

t = (80-fit60)/(sigma_hat*sqrt(1+1/n+((60-mean(mid))^2)/SSX))
pt(t, n-2, lower.tail = F)
```
Prediction interval for student who scored:
40 on midterm is (43.5125, 83.7253)
60 on midterm is (52.3891, 91.8589)
80 on midterm is (60.7394, 100.5186)

The probability that a student scores an 80 or above on the final when they scored a 60 on the midterm is 0.2113

