---
output:
  word_document: default
  html_document: default
  pdf_document: default
---
Problem 1.6.1
```{r}
table = scan('TABLE164.txt')
mean_table = mean(table)
t = qt(.1, length(table)-1)
low_bound = mean_table + t*sqrt(var(table))/sqrt(length(table))
high_bound = mean_table - t*sqrt(var(table))/sqrt(length(table))
mean_table
low_bound
high_bound
high_bound-low_bound
```

a) Sample mean = 6.989, Confidence Interval: 6.140 <= μ <= 7.838
b)C[6.140 <= μ <= 7.838] = .80
```{r}
t = qt(.05, length(table)-1)
low_bound = mean_table + t*sqrt(var(table))/sqrt(length(table))
low_bound
```
c) 95% lower confidence bound is 5.889
d) C[5.889<= μ] = .95
e) We are 95% confident that the true population mean lies above 5.889.
f)


g)NH: σ <= 5.0    AH: σ > 5.0     α=0.05
```{r}
chisq = 29*var(table)/25
p = pchisq(chisq, 29, lower.tail = F)
p
```
Since P>α, I do not reject the null hypothesis.



```{r}
chi_upper = qchisq(.95, 29)
chi_lower = qchisq(.05, 29)
upper = sqrt(29*var(table)/chi_upper)
lower = sqrt(29*var(table)/chi_lower)
upper
lower
```
h)C[2.925 <= σ <= 4.534] = .90. We can accept the null hypothesis at a 90% confidence level, since the 90% confidence interval is all less than 5



```{r}
t=qt(.995, length(table)-1)
low_bound = mean_table - t*sqrt(var(table))/sqrt(length(table))
high_bound = mean_table + t*sqrt(var(table))/sqrt(length(table))
low_bound
high_bound
high_bound-low_bound
```

i)interval is [5.205, 8.772]. Width is 3.566, width in a) was 1.697. Width is greater with more confidence. 




Problem 1.81
```{r}
A = matrix(c(9,4,3,4,16,8,3,8,12), nrow = 3, ncol = 3, byrow = TRUE)
A_transpose = t(A)
A_transpose
```
a) ^^^^^ See Above

b)C+B^T : Defined, same dimensions
  B+C : Undefined, different dimensions
  B+C^T : Defined, same dimensions
  AC : Defined, # of A columns == # of C rows
  CA : Undefined, # of C columns != # of A rows
  B-C^T : Defined, same dimensions
  
  
  
```{r}
A = matrix(c(9,4,3,4,16,8,3,8,12), nrow = 3, ncol = 3, byrow = TRUE)
B = matrix(c(12,14,3,4,31,5,5,13,21,6,2,31), nrow = 4, ncol = 3, byrow = TRUE)
C = matrix(c(12,23,17,22,24,28,19,20,31,30,41,27), nrow = 3, ncol = 4, byrow = TRUE)
print('C+B^T')
C + t(B)
print('B+C^T')
B+t(C)
print('AC')
A %*% C
print('B-C^T')
B - t(C)
```
c) See Above ^^



Problem 1.83
```{r}
x = matrix(c(1,2,1,3,1,1,1,4), nrow = 4, ncol = 2, byrow = TRUE)
y = matrix(c(1,6,3,2))
```
a)(X^T)= See Below
```{r}
t(x) %*% x
```

b) (X^T)y = See Below
```{r}
t(x) %*% y
```

c)((X^T)X)^-1 = See Below
```{r}
solve(t(x) %*% x)
```

d)$$X^TXB = X^Ty$$
```{r}
left = t(x) %*% x
right = t(x) %*% y
solve(left, right)
```
B/Beta = ^^


Problem 1.10.8
Lower confidence bound gives the appropriate information. We care about if the new strength is better than the old strength, the upper bound doesn't matter.


Problem 1.10.9
Null Hypothesis: μy <= 540, Alternative Hypothesis: μy > 540


Problem 1.10.10
α= 0.05
```{r}
C = qt(.95, 20-1)
Sample_mean = 12000/20
tc = (Sample_mean - 540)/(14/sqrt(20))
tc
C
```
We reject the null hypothesis since tc > C.


Problem 1.10.11
```{r}
pt(tc, 20-1, lower.tail = F)
```
P-value gives more reason to reject the NH. The probability of obtaining the sample mean or a value more extreme if the null hypothesis is true is almost impossible.




Exam Problem
```{r}
exam = read.table('EXAM.txt', header = T)
fin = exam$FIN
mid = exam$MID
```
Calculate 95% Confidence Interval
a)
```{r}
mid_mean = mean(mid)
mid_std = sqrt(var(mid))
z = qnorm(.975)
t = qt(.975, length(mid))
width = t*mid_std/sqrt(length(mid))
upperb = mid_mean + width
lowerb = mid_mean - width
upperb
lowerb
```



b)
```{r}
fin_mean = mean(fin)
fin_std = sqrt(var(fin))
z = qnorm(.975)
t = qt(.975, length(fin))
width = t*fin_std/sqrt(length(fin))
upperb = fin_mean + width
lowerb = fin_mean - width
upperb
lowerb
```

Hypothesis Test

Null Hypothesis: Fin Mean = Mid Mean
Alternative Hypothesis: Fin Mean =/= Mid Mean
α= 0.1
```{r}
t.test(fin, mid)
C = qt(.95, 44.551)
C
```
Since |t| is not greater than C, we do not have enough evidence to reject the null hypothesis.


Scatterplot
```{r}
plot(fin, mid, xlab = 'Final Scores', ylab = 'Mid Scores')
lines(fin, lm(mid~fin)$fitted)
```
Straight line regression seems justified. There appears to be a linear relationship midterm and final scores outside of the one outlier near mid:60, fin:90.