---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---

1)

4.4.1)
```{r}
table = read.table('table445.txt', header = T)
Y = table$Strength
ones = c(1,1,1,1,1,1,1,1,1,1)
X = matrix(data = c(ones,table$Temp, table$Pres), nrow = 10)
#colnames(x) = c("Temp", "Pres")
Y
X
```
^Y is the vector, X is the matrix

4.4.2)
```{r}
Xt = t(X)
XtX = Xt %*% X
XtY = Xt %*% Y
print('X^tX')
XtX
print('X^tY')
XtY
```
Matrices above^

4.4.3)
```{r}
XtX_inverse = solve(XtX)
beta = XtX_inverse %*% XtY
beta
```
beta hat^ These values for beta hat are the same as the ones in exhibit 4.4.4

4.4.4)
```{r}
SSY = sum((Y-mean(Y))^2)
x1 = X[,2]
x2 = X[,3]
SSX1 = sum((x1-mean(x1))^2)
SSX2 = sum((x2-mean(x2))^2)
SSY
SSX1
SSX2
```
SSY = 547.161
SSX1 = 7090
SSX2 = 115.6

4.4.6)
```{r}
e = Y-X%*%beta
SSE = t(e)%*%e
sigma = sqrt(SSE/(length(Y)-2-1))
sigma

```
sigma hat = 1.1169
population sigma is 1.7076
sigma hat underestimates sigma by about 0.59

4.4.7)
```{r}
estimate = beta0 + 300*beta1 + 16*beta2
population = -5 + 300*.2 + 16*-1
estimate
population
```

muy(300, 16) hat is 38.4715. The estimated value is fairly close to the population value, only underestimated by 0.53.

4.4.10)
```{r}
estimate = beta0 + 280*beta1 + 18*beta2
estimate
```
muy(280, 18) hat = 32.59673

4.4.12) 
```{r}
y_estimate = beta0 + 250*beta1 + 16*beta2
pnorm((31-y_estimate)/sigma, lower.tail = F)
```
About 4.54% of containers have a strength greater than 31 units if they are produced at a temperature of 250 degrees C and presure of 16 psi.




2)
```{r}
senic=read.table("http://www.utstat.toronto.edu/~brunner/data/legal/openSENIC.data.txt")
senic=senic[,-2]
senic=na.omit(senic)
```

```{r}
#beta1: nbeds
#beta2: nurses
#beta3: lngstay
#beta4: age
Y = senic$infpercent
ones = c(rep(1, nrow(senic)))
X = matrix(c(ones, senic$nbeds, senic$nurses, senic$lngstay, senic$age), nrow = nrow(senic))
Xt = t(X)
XtX = Xt %*% X
XtY = Xt %*% Y
XtX_inverse = solve(XtX)
beta = XtX_inverse %*% XtY
e = Y - X %*% beta
et = t(e)
SSE = et %*% e
sigma = sqrt(SSE/(nrow(senic)-4-1))
print('X^tX')
XtX
print('X^tY')
XtY
print('beta hat')
beta
print('sigma hat')
sigma
```
```{r}
sigma = sigma[1]
beta0_error = sigma*sqrt(XtX_inverse[1,1])
beta1_error = sigma*sqrt(XtX_inverse[2,2])
beta2_error = sigma*sqrt(XtX_inverse[3,3])
beta3_error = sigma*sqrt(XtX_inverse[4,4])
beta4_error = sigma*sqrt(XtX_inverse[5,5])
beta0_error
beta1_error
beta2_error
beta3_error
beta4_error
```
```{r}
lm1 = lm(infpercent~nbeds + nurses + lngstay + age, data = senic)
summary(lm1)
```
Ordering of variables
beta1: nbeds
beta2: nurses
beta3: lngstay
beta4: age

X^tX, X^tY, beta hat are above the error part ^^
sigma hat = 1.0732

se of beta0: 1.374
se of beta1: 0.001561002
se of beta2: 0.002132096
se of beta3: 0.06240142
se of beta4: 0.02600385

Results are the same as in lm()



2)
```{r}
sres = rstandard(lm1)
plot(senic$nbeds, sres,pch=20,xlab='nbeds',ylab='standardized residual', main='Rstandard versus nbeds')
plot(senic$lngstay, sres,pch=20,xlab='lngstay',ylab='standardized residual', main='Rstandard versus lngstay')
plot(senic$age, sres,pch=20,xlab='age',ylab='standardized residual', main='Rstandard versus age')
plot(senic$nurses, sres,pch=20,xlab='nurses',ylab='standardized residual', main='Rstandard versus nurses')

plot(lm1$fitted.values, sres,pch=20,xlab='fitted',ylab='standardized residual', main='Rstandard versus fitted')

boxplot(sres~senic$region, names = c('NorthCentral', 'NE', 'South', 'West'), xlab = 'Region')

qqnorm(sres,pch=1,xlab='Normalized Value',ylab='Standardized Residual',
main='QQ Plot of Standardized Residuals')
```
NE = NorthEast, couldn't fit on box plot
The model seems like a good fit for the data. The residuals appear to be fairly randomly distributed for each of the covariates and the fitted values. The QQ-plot of the residuals is linear. Each region has similar ranges of residual values.

3)
```{r}
df = data.frame(nbeds=531,nurses=442,lngstay=9.1,age=55)

a = as.numeric(c(1, df))
pred=as.numeric(t(a) %*% beta)
se=as.numeric(sigma*sqrt(t(a) %*% XtX_inverse %*% a))
pred+c(0,-1,1)*qt(0.995,nrow(senic)-4-1)* se
pse=as.numeric(sigma*sqrt(1+t(a) %*% XtX_inverse %*% a))
pred+c(0,-1,1)*qt(0.95,nrow(senic)-4-1)* pse

                
predict.lm(lm1, df, interval = 'confidence', level = 0.99)
predict.lm(lm1, df, interval = 'prediction', level = 0.90)
```
i) 99% confidence interval for mean infpercent with above parameters is (4.1717, 5.7870)
ii) 90% prediction interval for infpercent with above parameters is (3.1237, 6.8349)

We care more about the prediction interval. We want to estimate infpercent for this particular new hospital.