---
title: 'Homework 2: Data Frames and Apply'
output: pdf_document
---

Name:  Jeffrey Yang

Collaborated with:Tao Huang

This homework is due on Sept. 17th at 11:59pm. You can collaborate with your classmates, but you must identify their names above, and you must submit **your own** homework as a PDF file on Sakai.

**A special note**: You might see **Challenge** questions. You are not required to answer these challenge questions. However, 

- if you answered the challenge questions, you will get bonus points, allowing you to make mistakes in other parts of the homework and still get 10/10. 

- the maximum score is 10, so your score is capped at 10.  

You also see some optional questions. Again, you are not required to answer these questions, but additional credits will be given. 

**This week's agenda**: investigating the differences between data frames and matrices; practicing how to use the apply family of functions.

Crime data set
===

We're going to look at a data set containing the number of assaults, murders, and rapes per 100,000 residents, in each of the 50 US states in 1973. This comes from a built-in data frame called `USArrests`. We'll rename this to `crime.df` and append a column that gives the region for each state, from the built-in vector `state.region`. You can learn more about this crime data set by typing `?USArrests` into your R console.

```{r}
crime.df = data.frame(USArrests, Region = state.region)
```

Data frame basics
===

- **1a.** Report the number of rows of `crime.df`, and print its first 6 rows. Using the functions `is.data.frame()` and `is.matrix()`, confirm that it is a data frame, and not a matrix.

```{r}
#hit:
nrow(crime.df)
head(crime.df)
is.matrix(crime.df)
is.data.frame(crime.df)
```

- **1b.** We're ready to start investigating the differences between data frames and matrices. Use the `as.matrix()` function to convert `crime.df` into a matrix, calling the result `crime.mat`. Print the first 6 rows of `crime.mat`. Next, convert only the first 4 columns of `crime.df` into a matrix, and call the result `crime.mat.noregion`. Print the first 6 rows of `crime.mat.noregion`. Take a look at the first 6 rows of `crime.df`, `crime.mat`, and `crime.mat.noregion`. There is something unsatisfactory about `crime.mat`. What is it and why did this happen? (hit: matrix can only have one type of data.) If you need some guidance, try using the `class()` function to figure out the class of the first in each of the three objects.
```{r}
crime.mat = as.matrix(crime.df)
crime.mat.noregion = as.matrix(crime.df[,1:4])
head(crime.mat)
head(crime.mat.noregion)
```
Numbers in crime.mat were converted into strings because the region column is strings. The matrix must have all strings since matrix can only have 1 data type.


- **1c.** We now move to another difference between data frames and matrices, with regard to column access/indexing.  You can access the `Murder` column of `crime.df` by also typing in `crime.df$Murder`. Print out the result. Try using this same strategy to access `Murder` column of `crime.mat.noregion`. Describe the difference (if any) between the two results. Note: you will need to set `error=TRUE` as an option in this code chunk to allow R Markdown to knit your lab, despite the the error you will encounter here.
```{r error=TRUE}
crime.df$Murder
crime.mat.noregion$Murder
```
Using $ operator on matrix creates error, while we can use it for df.


- **1d.** Lastly, we'll demonstrate another difference between data frames and matrices, with regard to column additions. Compute a vector called `TotalCrime` of length 50 that gives the sum of the values in `Murder`, `Assault` and `Rape` for each of the 50 states. The first element of `TotalCrime` should give the total crime in Alabama, the second element should give that in Alaska, etc. Do not use a `for()` loop for this; use `rowSums()` instead. Now, insert `TotalCrime` as a column to `crime.df`, and make sure your new column is named `TotalCrime` in the data frame. Note: there are many ways to do this. Print the first 6 rows of the new `crime.df` data frame (hit: `head()`).
```{r}
TotalCrime = rowSums(crime.df[,c(1,2,4)])
crime.df = cbind(crime.df, TotalCrime)
head(crime.df)
```




Apply functions versus `for()` loops
===

The purpose of the next several questions is to help you internalize how the apply functions---specifically, `apply()`, `sapply()`, `lapply()`, and `tapply()`---are essentially convenient ways to write `for()` loops. 

Here's an example to get us started. Consider the following list, called `lis`, which contains 4 vectors of 5 randomly generated numbers.

```{r}
set.seed(10)
lis = list(rnorm(5), rnorm(5), rnorm(5), rnorm(5))
lis
```

Suppose we wanted to compute the mean of each vector (so we're looking for 4 numbers). We could do this using a `for()` loop in the following way, storing the results in `mean.vector`.

```{r}
mean.vector = vector(length=length(lis), mode="numeric")
for (i in 1:length(lis)) {
  mean.vector[i] = mean(lis[[i]])
}
mean.vector
```

We could also do this using a call to `sapply()`, in the following simpler way, storing the result as `mean.vector2`. This gives us the same exact answer.

```{r}
mean.vector2 = sapply(lis, mean)
all.equal(mean.vector, mean.vector2)
```

We're going to ask you to emulate this for each of 3 other apply functions (`lapply()`, `apply()` and `tapply()`) in the next 3 questions. Your goal will be to compute something using one of the apply functions or a `for()` loop, and show they are the same. The tricky part here will be formatting the `for()` loop properly to match exactly the apply function's output.
```{r}

```


- **2a.** Compute the standard deviation of each of the 4 vectors in `lis`, in two ways. For the first way use `lapply()`, in just one line of code, and call the result `sd.list`. For the second, use a `for()` loop, and call the result `sd.list2`. Use `all.equal()` to show that `sd.list` and `sd.list2` are the same. Hint: to construct an empty list of length `n`, you can use the command `vector(length=n, mode="list")`.

```{r}
# the easy way, using lapply
sd.list = lapply(lis, FUN=sd)
#create an empty list of the correct size then calculate the
# standard deviation of each vector in lis using a for loop
sd.list2 = vector(length=length(lis), mode="list")

for (i in 1:length(lis)) {
sd.list2[[i]] = sd(lis[[i]])
}

sd.list
all.equal(sd.list, sd.list2)
```

- **2b.** (optional) Using `crime.mat.noregion`, compute the maximum value in each of the 5 columns, in two ways. For the first way, use `apply()`, in just one line of code, and call the result `max.vector`. For the second, use a `for()` loop, and call the result `max.vector2`. Use `identical()` to show that `max.vector` and `max.vector2` are equal.

```{r}
# using apply first
max.vector = as.vector(apply(crime.mat.noregion, MARGIN = 2, FUN = max))

# create an empty vector of the correct dimensions
max.vector2 = vector(length = ncol(crime.mat.noregion),mode = "numeric")

# loop through, calculating the maximum value of each column
for(i in 1:ncol(crime.mat.noregion)) {
max.vector2[i] = max(crime.mat.noregion[,i])
}

max.vector
identical(max.vector,max.vector2)
```


- **2c.** (optional) Using `crime.df`, compute the minimum value of `Murder` within each of the four regions (Northeast, South, North Central, and West), in two ways. For the first way, use `tapply()`, in just one line of code, and call the result `min.vector`. For the second, use a `for()` loop, and call the result `min.vector2`. Show `min.vector` and `min.vector2` are equal. Hint: the trickiest part to figure out here is how to get the order of values in `min.vector` and `min.vector2` to be the same. Use `levels(crime.df$Region)` to dictate the order of regions in `min.vector2`. You'll also have to cast `min.vector2` to be the same data structure as `min.vector`.

```{r}
# using tapply
min.vector = as.vector(tapply(crime.df$Murder, INDEX = crime.df$Region, FUN = min))

# create an empty vector of the correct dimensions
min.vector2 = vector(length(levels(crime.df$Region)),mode = "numeric")

# save the region names now to make code readable
regions = levels(crime.df$Region) 

# loop through each region, calculating the minimum value in that region
for (i in 1:length(regions)) {
min.vector2[i] <- min(crime.df$Murder[crime.df[,"Region"] == regions[i]])

}
min.vector
# check the two lists are the identical 
all.equal(min.vector, min.vector2)
```
