---
title: 'Lab 2: R Basics'
date: "Week 3 (8/31/2021)"
output: pdf_document
---

Name: Jeffrey Yang
Collaborated with:Olivia, Anna, Alyssa

This lab is to be done in the lab session (completed outside of class if need be). You can collaborate with your classmates, but you must identify their names above, and you must submit **your own** lab as an knitted PDF/html file. The deadline for each lab session is

- **Deadline for 320.404: W 6:30 PM**
- **Deadline for 320.406: F 3:15 PM**
- **Deadline for 320.407: W 7:35 PM**

```{r}
## For reproducibility --- don't change this!
set.seed(01182018)
```

**This week's agenda**: manipulating data objects; using built-in functions, doing numerical calculations, and basic plots; introducing some probabilistic ideas.

**A special note**: You might see **Challenge** questions. You are not required to answer these challenge questions. However, 

- if you answered the challenge questions, you will get bonus points, allowing you to make mistakes in other parts of the lab and still get 10/10. 

- the maximum score is 10, so your score is capped at 10.  

You also get an example on how to answer lab questions for Question 1 in this lab. 

The binomial distribution
===

The binomial distribution $\mathrm{Bin}(m,p)$ is defined by the number of successes in $m$ independent trials, each have probability $p$ of success. Think of flipping a coin $m$ times, where the coin is weighted to have probability $p$ of landing on heads.

The R function `rbinom()` generates random variables with a binomial distribution. E.g., 

```{r,eval=FALSE}
rbinom(n=20, size=10, prob=0.5)
```

produces 20 observations from $\mathrm{Bin}(10,0.5)$.

Some simple manipulations
===

- **1a.** Generate 200 random values from the $\mathrm{Bin}(10,0.5)$ distribution, and store them in a vector called `bin.draws.0.5`. Try to: 1) extract and display the first 10 elements; 2) extract and display all but the first 175 elements.
```{r}
bin.draws.0.5 = rbinom(n=200, size=10, prob=0.5)
bin.draws.0.5[1:10]
bin.draws.0.5[-(1:175)]
```


- **1b.** Add the first element of `bin.draws.0.5` to the seventh. Compare the second element to the fifth, which is larger? A bit more tricky: print the indices of the elements of `bin.draws.0.5` that are equal to 5. How many such elements are there? **Challenge**: theoretically, how many such elements would you expect there to be?
```{r}
bin.draws.0.5[1] + bin.draws.0.5[7]
bin.draws.0.5[2] > bin.draws.0.5[5]
which(bin.draws.0.5==5)
length(which(bin.draws.0.5==5))
choose(10,5)*0.5^5*0.5^5*200 #expected number of elements
```


- **1c.** Find the mean and standard deviation of `bin.draws.0.5`. Is the mean close what you'd expect? The standard deviation?
```{r}
mean(bin.draws.0.5)
sd(bin.draws.0.5) #n*p*(1-p)

sqrt(10*0.5^2) #expected value;
```
Yes mean is close to expected value.

- **1d.** Call `summary()` on `bin.draws.0.5` and describe the result.
```{r}
summary(bin.draws.0.5)
```
The summary gives the minimum, 25% quantile, median, mean, 75% quantile, maximum value.

- **1e.** Find the data type of the elements in `bin.draws.0.5` using `typeof()`. Then convert `bin.draws.0.5` to a vector of characters (using the function `as.character()` ), storing the result as `bin.draws.0.5.char`, and use `typeof()` again to verify that you've done the conversion correctly. Call `summary()` on `bin.draws.0.5.char`. Is the result formatted differently from what you saw above? Why?
```{r}
typeof(bin.draws.0.5)
bin.draw.0.5.char=as.character(bin.draws.0.5)
typeof(bin.draw.0.5.char)
summary(bin.draw.0.5.char)
```
No, since a summary on on a vector characters is expected to be different that a summary of numbers.

Some simple plots
===

- **2a.** The function `plot()` is a generic function in R for the visual display of data. The function `hist()` specifically produces a histogram display. Use `hist()` to produce a histogram of your random draws from the binomial distribution, stored in `bin.draws.0.5`. 
```{r}
hist(bin.draws.0.5)
```


- **2b.** Call `tabulate()` on `bin.draws.0.5`. What is being shown? Does it roughly match the histogram you produced in the last question?
```{r}
tabulate(bin.draws.0.5)
```
Tabulate counts the number of times each number occurs in the vector. Looks about the same as the histogram. 

- **2c.** Call `plot()` on `bin.draws.0.5` to display your random values from the binomial distribution. Can you guess what the `plot()` function is doing here?
```{r}
plot(bin.draws.0.5)
```
Creates a scatterplot of the integer values, where the y axis is the value and the x axis is the index in the vector.


- **2d.** Call `plot()` with two arguments, the first being `1:200`, and the second being `bin.draws.0.5`. This creates a scatterplot of `bin.draws.0.5` (on the y-axis) versus the indices 1 through 200 (on the x-axis). Does this match your plot from the last question?
```{r}
plot(1:200, bin.draws.0.5)
```
Yes the scatterplot matches the one above, same indices.

More binomials, more plots
===

- **3a.** Generate 200 binomials again, but change the probability of success to: 0.2, 0.3, 0.4, 0.6, 0.7, and 0.8, storing the results in vectors called `bin.draws.0.2`, `bin.draws.0.3`, `bin.draws.0.4.`, `bin.draws.0.6`, `bin.draws.0.7` and  `bin.draws.0.8`. For each, compute the mean and standard deviation.
```{r}
bin.draws.0.2 = rbinom(n=200, size=10, prob=0.2)
bin.draws.0.3 = rbinom(n=200, size=10, prob=0.3)
bin.draws.0.4 = rbinom(n=200, size=10, prob=0.4)
bin.draws.0.6 = rbinom(n=200, size=10, prob=0.6)
bin.draws.0.7 = rbinom(n=200, size=10, prob=0.7)
bin.draws.0.8 = rbinom(n=200, size=10, prob=0.8)

mean(bin.draws.0.2)
sd(bin.draws.0.2)
mean(bin.draws.0.3)
sd(bin.draws.0.3)
mean(bin.draws.0.4)
sd(bin.draws.0.4)
mean(bin.draws.0.6)
sd(bin.draws.0.6)
mean(bin.draws.0.7)
sd(bin.draws.0.7)
mean(bin.draws.0.8)
sd(bin.draws.0.8)
```


- **3b.** We'd like to compare the properties of our vectors. Create a vector of length 7, whose entries are the means of the 7 vectors we've created, in an order according to the success probabilities of their underlying binomial distributions (0.2 through 0.8).
```{r}
vector = c(mean(bin.draws.0.2),mean(bin.draws.0.3),mean(bin.draws.0.4),mean(bin.draws.0.5),mean(bin.draws.0.6),mean(bin.draws.0.7),mean(bin.draws.0.8))
vector
```



Working with matrices
===

- **4a.** Create a matrix of dimension 200 x 7, called `bin.matrix`, whose columns contain the 7 vectors we've created, in an order of the success probabilities of their underlying binomial distributions (0.2 through 0.8). 
```{r}
all_values = c(bin.draws.0.2,bin.draws.0.3,bin.draws.0.4,bin.draws.0.5,bin.draws.0.6,bin.draws.0.7,bin.draws.0.8)
bin.matrix = matrix(data = all_values,nrow = 200, ncol = 7, byrow = FALSE)
```


- **4b.** Print the first three rows of `bin.matrix`. Print the element in the 77th row and 6th column. Compute the largest element in first column. Compute the largest element in all but the first column.
```{r}
bin.matrix[1:3,]
bin.matrix[77,6]
max(bin.matrix[,1])
max(bin.matrix[,2:7])
```

- **4c.** Calculate the means of each column by calling `colMeans()` on `bin.matrix`. 
```{r}
colMeans(bin.matrix)
```


- **4d.** Compare the means you computed in the last question to those you computed in Q3b. Do they appear equal, when you print the results out to the console? Try using `identical()` to check that they are exactly equal.
```{r}
identical(vector, colMeans(bin.matrix))
```



**Challenge** All the questions below are challenge questions.


Warm up is over, let's go big
===

- **5a.** R's capacity for data storage and computation is very large compared to what was available 10 years ago. Generate 2 million numbers from $\mathrm{Bin}(1 \times 10^6, 0.5)$ distribution and store them in a vector called `big.bin.draws`. Calculate the mean and standard deviation of this vector.
```{r}
big.bin.draws = rbinom(2000000, 10^6, 0.5)
mean(big.bin.draws)
sd(big.bin.draws)
```


- **5b.** Create a new vector, called `big.bin.draws.standardized`, which is given by taking `big.bin.draws`, subtracting off its mean, and then dividing by its standard deviation. Calculate the mean and standard deviation of `big.bin.draws.standardized`. (These should be 0 and 1, respectively, or very, very close to it; if not, you've made a mistake somewhere).
```{r}
big.bin.draws.standardized = (big.bin.draws-mean(big.bin.draws))/sd(big.bin.draws)
mean(big.bin.draws.standardized)
sd(big.bin.draws.standardized)
```


- **5c.** Plot a histogram of `big.bin.draws.standardized`. To increase the number of histogram bars, set the `breaks` argument in the `hist()` function (e.g., set `breaks=100`). What does the shape of this histogram appear to be? Is this surprising? What could explain this phenomenon? (Hint: rhymes with "Mental Gimmick Serum" ...)
```{r}
hist(big.bin.draws.standardized, breaks = 100)
```
Histogram looks like a bell curve. Not surprising since we used a binomial distribution with large sample.

- **5d.** Calculate the proportion of times that an element of `big.bin.draws.standardized` exceeds 1.644854. Is this close to 0.05? 
```{r}
length(which(big.bin.draws.standardized > 1.644854))/length(big.bin.draws.standardized)
```
Yes, it's close to 0.05.

- **5e.** Either by simulation, or via a built-in R function, compute the probability that a standard normal random variable exceeds 1.644854. Is this close to 0.05? (Hint: for either approach, it would be helpful to look at the help file for the `rnorm()` function.)
```{r}
pnorm(1.644854, lower.tail = FALSE)
```
Yes, it's close to 0.05.
