---
title: "R Notebook"
output:
  word_document: default
  html_notebook: default
---
PLEDGE:I will neither give nor receive unauthorized aid in this exam.
SIGNED: Jeffrey Yang

1) 

a)C
b)A
c)D
d)B
e)B
f)A
g)B
h)D
i)C


2)
a)
```{r}
t = qt(.975, 15-1)
upper = 48.7 + t*6.9/sqrt(15)
lower = 48.7 - t*6.9/sqrt(15)
upper
lower
```

C[44.8789 <= mu of y <= 52.5211] = 0.95


b)
NH: mu of x = 60       AH: mu of x < 60
```{r}
C = qt(.99, 19-1)
tc = (57.9-60)/(8.3/sqrt(19))
tc
C
```
Since Tc > -C, we do not have enough evidence to reject the null hypothesis. For now, we accept the null hypothesis. 


C)
```{r}
numerator = (15-1)*(6.9^2) + 19-1*(8.3^2)
denominator = 15+19-2
pooledSD = sqrt(numerator/denominator)
pooledSD
t = qt(.95, 15+19-2)
upper = (57.9-48.7) + t*pooledSD*sqrt((1/15)+(1/19))
lower = (57.9-48.7) - t*pooledSD*sqrt((1/15)+(1/19))
upper
lower
t
```
Common SD = 4.3897
C[6.6317 <= mu of x - mu of y <= 11.7683] = 0.90

D)
NH: mu of x = mu of y       AH: mu of x =/= mu of y
```{r}
t = (57.9 - 48.7)/(pooledSD*sqrt((1/15)+(1/19)))
pt(t, 15+19-2, lower.tail = F)*2
```
There is strong evidence against the null hypothesis. The p-value from the hypothesis test is very close to zero and the confidence interval of the differences in means does not contain 0 in it.


3)
a)
```{r}
table = read.table('EXAM2004.txt', header = T)
x = table$MT
y = table$FIN
xbar = mean(x)
ybar = mean(y)
SSX = sum((x-xbar)^2)
SSY = sum((y-ybar)^2)
SXY = sum((x-xbar)*(y-ybar))
xbar
ybar
SSX
SSY
SXY
```
xbar = 64.1428
ybar = 73.8857
SSX = 14456.29
SSY = 5629.543
SXY = 6147.571

b)
```{r}
#Beta1 = SXY/SSX
beta1 = SXY/SSX
#Beta0 = ybar - beta1*xbar
beta0 = ybar - beta1*xbar
n = length(y)
SSE = SSY - ((SXY)^2)/SSX
sigmahat = sqrt(SSE/(n-2))
beta1
beta0
sigmahat
```
beta1 hat = 0.4252
beta0 hat = 46.6088
sigmahat = 9.5588


C)
```{r}
sum(pnorm(90.5, beta0+beta1*x, sigmahat)-pnorm(79.5, beta0+beta1*x, sigmahat))
sum(x>=79.5 & x<=90.5)
total = 0
for (value in x){
  total = total + pnorm(90.5, beta0+beta1*value, sigmahat)-pnorm(79.5, beta0+beta1*value, sigmahat)
}
total
```
The predicted number of students scoring between 79.5 and 90.5 is 8.4831. The actual amount was 8. If we take a conservative estimate of the predicted to be around 8, then the regression accurately predicted the number of students. 
